Synthesizing Symmetric Lenses

### 2019-01-30 papers/SymmetricLenses/main.tex
More polishing and comments

%%%%%%%%%%%%%%% BCP stopped here

As an example of the partnership between \Expand and \GreedySynth,
consider the lens shown in Figure~\ref{fig:example_lens}. 
Inferring this lens requires \Expand to use a star-semiring axiom\bcp{are we
assuming readers know what that is?} to
rewrite the regular expression \lstinline{("\n" . emp_salary)$^*$} within
\lstinline{emp_salaries} into 
\lstinline{"" | (("\n" . emp_salary). ("\n" . emp_salary)$^*$)} 
so that \GreedySynth can find the appropriate lens.

Our strategy of communicating synthesizers gives us a way to
enumerate pairs of stochastic regular expressions of increasing
rank and to efficiently search through them, but it poses a
problem: when do we stop? We may have found a promising lens
between a pair of stochastic regular expressions, but a different 
pair we haven't yet discovered may give rise to an even better lens. 
Or, we may have found a promising lens between a given pair of
stochastic regular expressions, but the discovered lens is not 
acceptable to the users.

Our algorithm must resolve a three-way tension among the quality of
the inferred lens, the number of examples the user must provide to
eliminate unwanted lenses, and the amount of time it takes to return a
result.  For example, if the algorithm has already found the lens in
Figure~\ref{fig:example_lens}, we don't want to spend a lot of time
searching for an even better lens.

To resolve this tension, our algorithm uses heuristics to judge whether to
return the current best satisfying lens to the user or to pass the next set of
unrollings to \GreedySynth. The heuristics favor stopping if the current best
satisfying lens has relatively high likelihood, indicating the lens is very
promising. The heuristics also favor stopping if \Expand has delivered to
\GreedySynth all pairs of stochastic regular expressions at a given rank and
there is a large number of pairs at the next rank because searching through all
such pairs will take a long time. With this approach, the algorithm can quickly
respond with a satisfying lens with relatively high likelihood. If users are
unhappy with the result, they can either refine the search by supplying
additional examples, which serve both to rule out the previously proposed
lens(es) and to reduce the size of the search space by cutting down on the
number of satisfying lenses, or they can supply annotations (discussed in
\S\ref{subsec:optimizations}) informing describing whether certain information
should be retained through format conversion.

@@

\section{Stochastic Regular Expressions}
\label{sec:sre}
To characterize the likely lenses, we must compute the expected number of bits
needed to recover a string in one data source from a synchronized string in the
other data source. To do this, we must first develop a probabilistic language
model for our language. We do this with \emph{stochastic regular expressions}~\cite{stoch-rnn,stoch-def},
regular expressions annotated with probability information. With this, we can
jointly express a language and a probability distribution over that language. In
the following grammar, $\String$ ranges over strings, and $\Probability$ ranges
over real numbers between 0 and 1, exclusive.
$$\Regex{},\RegexAlt{} \GEq{}  \String  \; | \; \emptyset \; | \; \PRegexStar{\Regex}{p} \; | \; \RegexConcat{\Regex_1}{\Regex_2} \; | \; \PRegexOr{\Regex_1}{\Regex_2}{p}$$

In the stochastic regular expression
$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation $\Probability$
represents the likelihood that the string is in $\Regex_1$. In the stochastic regular
expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
represents how
much more likely having a string involving more than $n$ iterations is than
having a string involving exactly $n$ iterations. The semantics of SREs is
defined by the function $P$ below.

\begin{center}
  \begin{tabular}{rcl}
    $\ProbabilityOf{\String}{\String''}$
    & =
    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\

@@

    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
    & =
    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n(1-\Probability)\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
  \end{tabular}
\end{center}

@@

\subsection{Stochastic Regular Expression Equivalences}
$\Expand$ needs to be able to find equivalent stochastic regular expressions.
However, existing work does not provide a means of reasoning about stochastic
regular expression equivalence. We extend the
\emph{star-semiring}~\cite{Droste2009} equivalences to stochastic regular
expressions (Figure~\ref{fig:ssr-sre}).

>>>>

To illustrate the partnership between \Expand and \GreedySynth,
consider the lens shown in Figure~\ref{fig:example_lens}. 
Inferring this lens requires \Expand to use a star-semiring axiom\bcp{are we
assuming readers know what that is?} to
rewrite the regular expression \lstinline{("\n" . emp_salary)$^*$} within
\lstinline{emp_salaries} into 
\lstinline{"" | (("\n" . emp_salary). ("\n" . emp_salary)$^*$)} 
so that \GreedySynth can find the appropriate lens.  \bcp{This example would
be more helpful if it were spelled out in full.  What is the path from the
start of the search to the point where we find this solution?}

This architecture of communicating synthesizers gives us a way to
enumerate pairs of stochastic regular expressions of increasing
rank and to efficiently search through them, but it poses a
problem: when do we stop?\bcp{This question would arise independent of the
  architecture of the search algorithm!} We may have found a promising lens
between a pair of stochastic regular expressions, but a different 
pair we haven't yet discovered may give rise to an even better lens. 
Or, we may have found a promising lens between a given pair of
stochastic regular expressions, but the discovered lens is not 
acceptable to the users. \bcp{How would the algorithm know this?}

The search algorithm must resolve a three-way tension among the quality of
the inferred lens, the number of examples the user must provide to
eliminate unwanted lenses, and the amount of time it takes to return a
result.  For example, if the algorithm has already found the lens in
Figure~\ref{fig:example_lens}, we don't want to spend a lot of time
searching for an even better lens.\bcp{How does this example illustrate a
  three-way tension?}

To resolve this tension, the algorithm uses heuristics to judge whether to
return the current best satisfying lens to the user or to pass the next set of
unrollings to \GreedySynth. The heuristics favor stopping if the current best
satisfying lens has relatively high\bcp{high compared to what??} likelihood,
indicating the lens is very 
promising. The heuristics also favor stopping if \Expand has delivered to
\GreedySynth all pairs of stochastic regular expressions at a given rank and
there is a large number of pairs at the next rank, because searching through all
such pairs will take a long time. With this approach, the algorithm can quickly
respond with a satisfying lens with relatively high likelihood. If the user is
unhappy with the result, they can either refine the search by supplying
additional examples, which serve both to rule out the previously proposed
lens(es) and to reduce the size of the search space by cutting down on the
number of satisfying lenses, or they can supply annotations on the source
and target SREs indicating that certain information
either must be retained or must be discarded by the lens (see
\S\ref{subsec:optimizations}\bcp{this reference to the later section with
  details is nice, but why don't we have similar references throughout this
  overview??}).

@@

\section{Stochastic Regular Expressions}
\label{sec:sre}
To characterize likely lenses, we must compute the expected number of bits
needed to recover a string in one data source from a synchronized\bcp{again,
needs a definition} string in the
other data source. To do this, we must first develop a probabilistic 
model for our language. We do this with \emph{stochastic regular
  expressions} (SREs)---regular expressions annotated with probability
information~\cite{stoch-rnn,stoch-def}---which 
jointly express a language and a probability distribution over that language. 
$$\Regex{},\RegexAlt{} \quad\GEq{}\quad  \String  \quad|\quad \emptyset \quad|\quad \PRegexStar{\Regex}{p} \quad|\quad \RegexConcat{\Regex_1}{\Regex_2} \quad|\quad \PRegexOr{\Regex_1}{\Regex_2}{p}$$
(Lowercase $\String$ ranges over constant strings and $\Probability$ ranges
over real numbers between 0 and 1, exclusive.)

The semantics of an SRE $S$ is a probability distribution $P_S$ defined as
follows.
\begin{center}
  \begin{tabular}{rcl}
    $\ProbabilityOf{\String}{\String''}$
    & =
    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\

@@

    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
    & =
    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n(1-\Probability)\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
  \end{tabular}
\end{center}
In the stochastic regular expression\bcp{abbreviate to SRE here and in many
  places elsewhere}
$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation $\Probability$
represents the likelihood that the string is in $\Regex_1$\bcp{That seems
  both confusing (what is ``the string''??) and probably inaccurate (what if
  $S_1$ and $S_2$ are the same?)}. In the stochastic regular
expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
represents how
much more likely having a string involving more than $n$ iterations is than
having a string involving exactly $n$ iterations. \bcp{Last sentence needs
  an example.  Probably the first sentence also.}

@@

\subsection{Stochastic Regular Expression Equivalences}
The $\Expand$ algorithm needs to be able to enumerate SREs that are
``equivalent'' to a given one.
However, existing work does not provide a means of reasoning about stochastic
regular expression equivalence\bcp{Does not provide a means of {\em
    reasoning} about it, or does not even define it?}. We extend the
\emph{star-semiring}~\cite{Droste2009} equivalences\bcp{again, we seem to be
assuming that all readers will already know what they are...} to stochastic regular
expressions (Figure~\ref{fig:ssr-sre}).
